---
title: "Use case: the Big Data PE API"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Use case: the Big Data PE API}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

`apifetch` was extracted from the
[BigDataPE](https://github.com/leite/BigDataPE) package, which wraps the
**Big Data PE** platform — a public-data REST API run by the Government of the
State of Pernambuco, Brazil. This vignette shows how that service maps onto a
single `apifetch` profile, and is the canonical worked example of configuring
the package for a real API.

## What is specific about Big Data PE

The Big Data PE API has two conventions that differ from the `apifetch`
defaults:

- **Authentication** sends the token *verbatim* in the `Authorization` header
  (no `Bearer ` prefix) — use `af_auth_raw()`.
- **Pagination** sends `limit`/`offset` as **HTTP headers**, not query
  parameters — use `af_paginate_offset("header")`.

It also returns a status column named `"Mensagem"` that we drop when combining
chunks (`drop_cols = "Mensagem"`), and it is only reachable from the
**PE Conectado** network or VPN, which we surface via `connect_hint`.

## Defining the profile

```{r}
library(apifetch)

bigdatape <- af_api(
  endpoint     = "https://www.bigdata.pe.gov.br/api/buscar",
  service      = "BigDataPE",
  auth         = af_auth_raw(),
  pagination   = af_paginate_offset("header"),
  drop_cols    = "Mensagem",
  connect_hint = "Ensure you are on the PE Conectado network or VPN."
)
```

## Storing a token and fetching

```{r}
af_store_token("dengue", "your-token-here", service = "BigDataPE")

# A single page of 50 records
dengue <- af_fetch(bigdatape, "dengue", limit = 50)

# Everything, in chunks, with a progress message per chunk
dengue_all <- af_fetch_all(
  bigdatape, "dengue",
  chunk_size = 50000,
  verbosity = 1
)
```

## Note on language

Function and parameter names are English (R convention), but the API's response
column names and some values are Portuguese (e.g. `nu_notificacao`,
`"BOA VIAGEM"`). That is intentional — they come straight from the upstream
service.
